Predicting tree failure likelihood for utility risk mitigation via artificial intelligence

### 2021-03-05 docs/manuscript/manuscript.tex
added appendix tables

%AMS-TeX packages
%\usepackage{amssymb,amsmath,amsthm}
%\usepackage{bm}
%\usepackage[mathscr]{eucal}
%\usepackage{colortbl}

@@

\subsection{Sensitivity to training resolution}
\autoref{fig:sens-val-acc} shows the boxplots of the validation performance metrics for all twelve cases considered.
The metrics are: accuracy and the macro-averages of precision, recall and the $F_{1}$ score.
%With respect to the accuracy, there are no significant differences between the 
Taking all metrics into consideration, Welch pairwise tests showed that there are no significant differences in performance among the 3 training resolutions, with one exception.
In the scenario Pr\_PoIm, the accuracy, $F_{1}^{m}$, and $Re^{m}$ for the 128-pixel case are greater than those for the 224-pixel case ($p<.001$).
This difference in performance is not as stark between the 128-pixel and 64-pixel cases in the same scenario.
This outcome implies that we can achieve efficiency by  training at lower resolutions without significant losses in performance.

% We further summarize the statistics (mean and standard deviation) of the validation accuracies in \autoref{tab:meanvalacc}.
% Using the Welch test for pairwise comparisons of resolution performance within each scenario, we found no statistically significant difference between the validation accuracy for each resolution ($p\ge 0.57$).

@@

\subsection{Scenario sensitivity}
We conducted pairwise tests between each scenario combination for all resolutions.
The results indicated that considering all metrics, the scenarios are statistically significantly different from each other ($p < .005$) except Pr\_Im compared to PrPo\_Im and Pr\_PoIm compared to Pr\_Po\_Im, with respect to certain metrics.
If the accuracy is not taken into account, then there is no strong evidence that Pr\_Im differs in performance when compared with PrPo\_Im.
In the case of Pr\_PoIm versus Pr\_Po\_Im, there is greater evidence to support their similarity in performance 
with respect to  $F_{1}^{m}$ and $Re^{m}$. %and $Pr^{m}$
Otherwise, the performance between these two scenarios is significantly different.
From \autoref{fig:sens-val-acc}, we observe the best performances in Pr\_Im and PrPo\_Im, with all metrics greater than or equal to 0.90.
The performance in Pr\_PoIm is lower, and Pr\_Po\_Im has the lowest performance of all four.

@@

\appendix

>>>>

%AMS-TeX packages
%\usepackage{amssymb,amsmath,amsthm}
\usepackage{bm}
%\usepackage[mathscr]{eucal}
%\usepackage{colortbl}

@@

\subsection{Sensitivity to training resolution}
\autoref{fig:sens-val-acc} shows the boxplots of the validation performance metrics for all twelve cases considered.
The metrics are: accuracy and the macro-averages of precision, recall and the $F_{1}$ score.
For reference, we summarize the statistics (mean and standard deviation) of the validation performance metrics in \autoref{tab:meanstd}.
%With respect to the accuracy, there are no significant differences between the 
Taking all metrics into consideration, Welch's pairwise tests show that there are no significant differences in performance among the 3 training resolutions, with one exception.
In the scenario Pr\_PoIm, the accuracy, $F_{1}^{m}$, and $Re^{m}$ for the 128-pixel case are greater than those for the 224-pixel case ($p<.001$).
This difference in performance is not as stark between the 128-pixel and 64-pixel cases in the same scenario.
This outcome implies that we can achieve efficiency by  training at lower resolutions without significant losses in performance.
(A complete summary of the p-values is provided in \autoref{tab:welch-resolution}.)
% Using the Welch test for pairwise comparisons of resolution performance within each scenario, we found no statistically significant difference between the validation accuracy for each resolution ($p\ge 0.57$).

@@

\subsection{Scenario sensitivity}
We conducted pairwise tests between each scenario combination for all resolutions. (A complete summary of the results is given in \autoref{tab:welch-scenario}.
The results indicated that considering all metrics, the scenarios are statistically significantly different from each other ($p < .005$) except Pr\_Im compared to PrPo\_Im and Pr\_PoIm compared to Pr\_Po\_Im, with respect to certain metrics.
If the accuracy is not taken into account, then there is no strong evidence that Pr\_Im differs in performance when compared with PrPo\_Im.
In the case of Pr\_PoIm versus Pr\_Po\_Im, there is greater evidence to support their similarity in performance 
with respect to  $F_{1}^{m}$ and $Re^{m}$. %and $Pr^{m}$
Otherwise, the performance between these two scenarios is significantly different.
From \autoref{fig:sens-val-acc}, we observe the best performances in Pr\_Im and PrPo\_Im, with all metrics greater than or equal to 0.90.
The performance in Pr\_PoIm is lower, and Pr\_Po\_Im has the lowest performance of all four.

@@

\appendix
\section{Summary of validation performance metrics}
For further reference we tabulate the mean and standard deviation of the performance metrics obtained on the scenario-resolution sensitivity tests. These values are shown in \autoref{tab:meanstd}.
The metrics are accuracy, macro-average $F_1$ ($F_1^m$), macro-average recall ($Re^m$) and macro-average precision ($Pr^m$).

\begin{table}[ht]\small\centering
\caption{Mean and standard deviation of the 5-fold cross-validation estimates of classifier performance metrics on the 12 scenario-resolution cases investigated. ``SD'' represents ``standard deviation.''}
\label{tab:meanstd}
\begin{tabular}{llllllllll}\toprule
\multirow{3}{*}{\textbf{Scenario}} &  
\multirow{3}{*}{\textbf{Resolution}}  & \multicolumn{2}{c}{\textbf{Accuracy}} & \multicolumn{2}{c}{$\bm{F_1^m}$} 
& \multicolumn{2}{c}{$\bm{Re^m}$} & \multicolumn{2}{c}{$\bm{Pr^m}$} \\%\cmidrule{3-10}
                          &            & Mean & SD  & Mean & SD  & Mean & SD  & Mean & SD  \\\midrule
\multirow{3}{*}{Pr\_Im}    & 64        & 0.94 & 0.01 & 0.92 & 0.02 & 0.9  & 0.02 & 0.94 & 0.01 \\
                          & 128        & 0.95 & 0.01 & 0.93 & 0.02 & 0.91 & 0.03 & 0.95 & 0.01 \\
                          & 224        & 0.94 & 0.01 & 0.92 & 0.01 & 0.91 & 0.01 & 0.93 & 0.02 \\\midrule
\multirow{3}{*}{PrPo\_Im}  & 64         & 0.92 & 0.01 & 0.91 & 0.01 & 0.91 & 0.02 & 0.92 & 0.01 \\
                          & 128        & 0.92 & 0.01 & 0.91 & 0.01 & 0.9  & 0.01 & 0.92 & 0.01 \\
                          & 224        & 0.91 & 0.01 & 0.91 & 0.01 & 0.9  & 0.01 & 0.91 & 0.01 \\\midrule
\multirow{3}{*}{Pr\_PoIm}  & 64         & 0.88 & 0.01 & 0.8  & 0.04 & 0.79 & 0.07 & 0.85 & 0.05 \\
                          & 128        & 0.91 & 0.01 & 0.85 & 0.02 & 0.84 & 0.02 & 0.87 & 0.03 \\
                          & 224        & 0.88 & 0.01 & 0.79 & 0.02 & 0.76 & 0.02 & 0.84 & 0.03 \\\midrule
\multirow{3}{*}{Pr\_Po\_Im} & 64         & 0.82 & 0.03 & 0.74 & 0.03 & 0.73 & 0.02 & 0.75 & 0.04 \\
                          & 128        & 0.84 & 0.02 & 0.76 & 0.03 & 0.74 & 0.03 & 0.77 & 0.03 \\
                          & 224        & 0.83 & 0.01 & 0.74 & 0.02 & 0.72 & 0.03 & 0.76 & 0.01\\ \bottomrule
\end{tabular}
\end{table}

\section{Welch $t$-test results}
Here, we summarize the p-values resulting from the Welch independent $t$-tests used to determine if the mean values of the various performance metrics are statistically significant or not.
Strong evidence of significant difference is given by a low p-value (a typical threshold is $< 0.01$).

\begin{table}[ht]\small\centering
\caption{Welch's $t$-test p-value results for pairwise comparisons between image resolutions across all four scenarios.
Low p-values indicates statistically different mean values for the validation metrics considered.}
\label{tab:welch-resolution}
\sisetup{scientific-notation=true, round-precision = 3,round-mode=figures,table-format = 1.2e-1,
exponent-product=\ensuremath{\cdot}}%output-exponent-marker = \text{e}, 
\begin{tabular}{@{}lll SSSS@{}}
\toprule
\multirow{2}{*}{\textbf{Scenario}} &
\multirow{2}{*}{\textbf{Resolution 1}} &
\multirow{2}{*}{\textbf{Resolution 2}} &
\multicolumn{4}{c}{\textbf{p-values}} \\ %\cmidrule(l){4-7} 
    & & & 
  \multicolumn{1}{c}{\textbf{Accuracy}} &
  \multicolumn{1}{c}{$\bm{F_1^m}$} &
  \multicolumn{1}{c}{$\bm{Re^m}$} &
  \multicolumn{1}{r}{$\bm{Pr^m}$} \\ \midrule
\multirow{3}{*}{Pr\_Im}    & 64  & 128 & 0.427766 & 0.472000 & 0.638283 & 0.243409 \\
                          & 64  & 224 & 0.810526 & 0.957770 & 0.728743 & 0.400994 \\
                          & 128 & 224 & 0.333005 & 0.421449 & 0.781772 & 0.119129 \\\midrule
\multirow{3}{*}{PrPo\_Im}  & 64  & 128 & 0.343214 & 0.334622 & 0.280111 & 0.607451 \\
                          & 64  & 224 & 0.112543 & 0.153209 & 0.299562 & 0.131835 \\
                          & 128 & 224 & 0.465132 & 0.566733 & 0.978346 & 0.253573 \\\midrule
\multirow{3}{*}{Pr\_PoIm}  & 64  & 128 & 0.006932 & 0.037836 & 0.139458 & 0.345129 \\
                          & 64  & 224 & 0.836653 & 0.715576 & 0.494841 & 0.697297 \\
                          & 128 & 224 & 0.001099 & 0.000207 & 0.000467 & 0.076234 \\\midrule
\multirow{3}{*}{Pr\_Po\_Im} & 64  & 128 & 0.241587 & 0.249616 & 0.391389 & 0.311935 \\
                          & 64  & 224 & 0.677163 & 0.937617 & 0.598361 & 0.650331 \\
                          & 128 & 224 & 0.227783 & 0.258068 & 0.252791 & 0.366722 \\ \bottomrule
\end{tabular}
\end{table}

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage{multirow}
\begin{table}[ht]\small\centering
\caption{Welch's $t$-test p-value results for pairwise comparisons between scenarios resolutions for all three image resolutions.
Low p-values indicates statistically different mean values for the validation metrics considered.}
\label{tab:welch-scenario}
\sisetup{scientific-notation=true, round-precision = 3,round-mode=figures,table-format = 1.2e-1,
exponent-product=\ensuremath{\cdot}}%output-exponent-marker = \text{e},

\begin{tabular}{@{}lllSSSS@{}}
\toprule
\multirow{2}{*}{\textbf{Scenario}} &
\multirow{2}{*}{\textbf{Resolution 1}} &
\multirow{2}{*}{\textbf{Resolution 2}} &
\multicolumn{4}{c}{\textbf{p-values}} \\ %\cmidrule(l){4-7} 
    & & & 
  \multicolumn{1}{c}{\textbf{Accuracy}} &
  \multicolumn{1}{c}{$\bm{F_1^m}$} &
  \multicolumn{1}{c}{$\bm{Re^m}$} &
  \multicolumn{1}{r}{$\bm{Pr^m}$} \\ \midrule
\multirow{6}{*}{64}  & Pr\_Im   & PrPo\_Im  & 0.005913 & 0.714479 & 0.422914 & 0.010639 \\
                     & Pr\_Im   & Pr\_PoIm  & 0.000068 & 0.001772 & 0.014794 & 0.008354 \\
                     & Pr\_Im   & Pr\_Po\_Im & 0.000184 & 0.000004 & 0.000004 & 0.000193 \\
                     & PrPo\_Im & Pr\_PoIm  & 0.000814 & 0.002534 & 0.010824 & 0.020167 \\
                     & PrPo\_Im & Pr\_Po\_Im & 0.000553 & 0.000014 & 0.000000 & 0.000351 \\
                     & Pr\_PoIm & Pr\_Po\_Im & 0.005380 & 0.029135 & 0.132204 & 0.007418 \\\midrule
\multirow{6}{*}{128} & Pr\_Im   & PrPo\_Im  & 0.004705 & 0.137997 & 0.578580 & 0.002389 \\
                     & Pr\_Im   & Pr\_PoIm  & 0.001249 & 0.000455 & 0.003295 & 0.000790 \\
                     & Pr\_Im   & Pr\_Po\_Im & 0.000005 & 0.000004 & 0.000012 & 0.000060 \\
                     & PrPo\_Im & Pr\_PoIm  & 0.169503 & 0.000446 & 0.001725 & 0.011313 \\
                     & PrPo\_Im & Pr\_Po\_Im & 0.000099 & 0.000064 & 0.000050 & 0.000320 \\
                     & Pr\_PoIm & Pr\_Po\_Im & 0.000132 & 0.000244 & 0.000244 & 0.000931 \\\midrule
\multirow{6}{*}{224} & Pr\_Im   & PrPo\_Im  & 0.000716 & 0.088647 & 0.576529 & 0.055010 \\
                     & Pr\_Im   & Pr\_PoIm  & 0.000002 & 0.000001 & 0.000025 & 0.000256 \\
                     & Pr\_Im   & Pr\_Po\_Im & 0.000000 & 0.000005 & 0.000029 & 0.000000 \\
                     & PrPo\_Im & Pr\_PoIm  & 0.000130 & 0.000003 & 0.000014 & 0.001828 \\
                     & PrPo\_Im & Pr\_Po\_Im & 0.000003 & 0.000014 & 0.000017 & 0.000000 \\
                     & Pr\_PoIm & Pr\_Po\_Im & 0.000144 & 0.003617 & 0.029345 & 0.001147 \\ \bottomrule
\end{tabular}
\end{table}
